---
title: "R Notebook"
output: html_notebook
---

# Coursera Practise I

## Import required libraries

```{r load-packages, message = FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(readxl)
library(lubridate)
library(ggplot2)
```

## Import Dataset and view first lines

```{r load data, results='hide'}
pm2_5 <- read_csv("./daily_SPEC_2014.csv.bz2")
```

## View data
Then, I'd like to view the `pm2_5` data
```{r view pm2_5}
head(pm2_5)
```

## Question I
What is average Sample.Value for "Bromine PM2.5 LC" in the state of Wisconsin in this dataset?

```{r average sample value}
Q1 <- pm2_5 %>% 
  group_by(`State Code`) %>% 
  filter(`Parameter Name` == "Bromine PM2.5 LC", `State Name` == "Wisconsin") %>% 
  summarise( average = mean(`Arithmetic Mean`, na.rm = TRUE))
```

Then, we could have a view on Q1 data

```{r view Q1}
head(Q1)
```

## Question II
Calculate the average of each chemical constituent across all states, monitoring sites and all time points.

```{r average}
Q2 <- pm2_5 %>% 
  group_by(`Parameter Name`, `State Name`, `Site Num`, `Date Local`) %>%  # Group data with these columns
  filter(grepl("LC", `Parameter Name`)) %>%                               # Find out all `Parameter Name` 
                                                                          # contains LC
  summarise( total_average = mean(`Arithmetic Mean`) ) %>%                # Create a column called total_average 
                                                                          #and store mean values into it
  arrange(desc(total_average))                                            # Arrange the dataset in descending order
```

Then, I would like to view the first lines of Q2

```{r view Q2}
head(Q2)
```

For better observing the result, I would like to plot the results out.

```{r plot Q2}
Q2_subtable <- Q2 %>%
  filter(grepl("^W", `State Name`))                 # Make a subset of Q2 by all the states start with `W`
ggplot(Q2_subtable, aes(x = `State Name`, y = total_average)) + 
  geom_bar(stat = "identity", colour = "red")       # Bar chart
```

## Question III
Which monitoring site has the highest average level of "Sulfate PM2.5 LC" across all time?

```

```
























