---
title: "My Practise I"
output: html_notebook
---

# Introduction
This file is mainly for practising how to use `dplyr` and `tidyr` package, in this practise, I would like to use the `hflights` data package.

## Load Packages
```{r load packages}
library(hflights)
library(tidyr)
library(dplyr)
library(ggplot2)
```

## View Data
Then, I'd like to view the `hflights` data

```{r view hflights}
head(hflights, 15)
```

Besides, I would like to know the total rows of `hflights`

```{r dim hflights}
dim(hflights)
nrow(hflights)
```

So there are 227496 data rows in the hflight.

For a better showing, the function tbl_df() could also be used.

```{r tbl_df}
tbl_hflights <- tbl_df(hflights)
```

## Practise I

Firstly, I would like to select the data of `FlightNum` equals to 428.

```{r filter}
flight_428 <- hflights %>%
  select(Year, Month, DayofMonth, FlightNum, Distance, Cancelled) %>%
  filter(FlightNum == 428)
```

Now, we could have a look at the `flight_428`

```{r view flight_428}
head(flight_428, 20)
dim(flight_428)
```

```{r plot flight_428}
ggplot(flight_428, aes(x = flight_428$Month, y = flight_428$Distance)) +
  geom_histogram(stat = "identity", fill = "red", alpha = 0.5) +
  labs(x = "Month", y = "Distance", title = "Fly Distance in Different Months of Flight 428") + 
  scale_x_continuous(breaks = seq(1, 12, 1))
```

From the graph, Janurary is the busiest month of Flight 428. Then, I would like to do a comparison.

```{r flight analysis}
flight_analysis <- hflights %>%
  select(Year, Month, DayofMonth, FlightNum, Distance) %>%
  group_by(Month) %>%
  summarise(month_total = n()) %>%
  mutate(month_total = month_total - 5254)                    # The original point start from 5254
```

```{r head flight_analysis}
head(flight_analysis)
```

Then, I would like to plot out the results, and observe the total flight numbers of each month

```{r plot flight_analysis}
ggplot(flight_analysis, aes(x = flight_analysis$Month, y = flight_analysis$month_total)) +
  geom_histogram(fill = "red", alpha = 0.5, stat = "identity") +
  geom_point(colour = "blue", alpha = 0.5, size = 2, y = flight_analysis$month_total / 2) +
  geom_line(colour = "blue", alpha = 0.5, size = 1, y = flight_analysis$month_total / 2) + 
  labs(x = "Month", y = "Month Total Flights") + 
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  geom_text(aes(label = flight_analysis$month_total, vjust = 0.5, hjust = 3, angle = 90), show_guide = FALSE)
```



















